\documentclass[11pt, letterpaper]{article}
\usepackage[margin=1in]{geometry}
\usepackage{enumitem}
\usepackage{parskip}
\usepackage{titlesec}
\usepackage[hidelinks]{hyperref}

\titleformat{\section}{\normalfont\bfseries}{}{0em}{}[\titlerule]
\titlespacing{\section}{0pt}{8pt}{4pt}

\begin{document}

\begin{center}
    {\Large \textbf{Micromancer's Realm}}\\[2pt]
    {\small CIS 5680 Group Game Project 1 --- Spring 2026}
\end{center}

\vspace{-4pt}

\section*{High Concept}

A botched alchemist's experiment shrinks you to the size of a thumbtack inside a craft shop---where
scissors are skyscrapers, spools of thread are canyon walls, and a spreading growth salve is about
to blow the whole place up. Switch between three tiny specialists and escape before the explosion.

\section*{Genre}

2D platformer / adventure with emergent environmental puzzles.

\section*{Main Features}

\begin{itemize}[leftmargin=1.5em, itemsep=3pt]

    \item \textbf{Tiny vs.\ Huge Scale.} Every everyday object in the craft shop is now a massive
    environmental feature. A toppled ruler becomes a bridge hundreds of meters long; a rolling
    pencil is an oncoming boulder; a puddle of paint is an impassable lake. The core sensation is
    the physical awe of being utterly dwarfed by the mundane---and finding clever ways through it.

    \item \textbf{Three Specialist Characters.} Players switch freely between three tiny heroes,
    each built for a different kind of obstacle. The \textbf{Acrobat} double-jumps and wall-kicks
    across yawning gaps, running along suspended thread. The \textbf{Digger} tunnels through soft
    obstacles---crumbling paper mache, glitter dunes---and shoves heavy items to clear paths. The
    \textbf{Engineer} builds temporary bridges from tape scraps and strings paperclip ziplines
    between anchor points. No single character can reach the exit alone.

    \item \textbf{Emergent Environment.} The growth salve continues to spread each level, randomly
    enlarging new objects as it goes. An avalanche of buttons may form a convenient ramp---until a
    cascading bolt of fabric knocks it flat. Players must read the shifting environment and adapt
    their character-switching strategy on the fly rather than execute a memorized solution.

    \item \textbf{Escalating Urgency.} A visible salve-spread timer counts down toward a volatile
    explosion. Early in each level the environment is relatively stable; as the timer ticks down,
    new growth events accelerate, paths close off, and the margin for error shrinks. Players who
    explore and solve efficiently are rewarded; dawdling is punished.

    \item \textbf{Physical Sensation Aesthetic.} The game leans into the tactile feel of miniature
    traversal: the bounce of landing on a stretched canvas, the slick slide across a wax crayon
    surface, the vibration of a giant spool spinning underfoot. Sound and screen-shake underscore
    scale at every moment.

\end{itemize}

\section*{Player Challenge / Motivation}

Players must read a dynamic, ever-changing oversized environment, decide which specialist to use for
each obstacle, and chain their abilities together to carve a path to the exit---all against an
expanding salve that constantly reshapes the level around them. The challenge is resourceful
improvisation under mounting pressure, not rote puzzle memorization.

\section*{Design Goals}

\begin{itemize}[leftmargin=1.5em, itemsep=2pt]
    \item \textbf{Physical Immersion.} Every interaction should sell the sensation of being tiny
    in a giant world---oversized objects have weight, inertia, and presence that the player
    must respect and exploit.
    \item \textbf{Emergent Strategy.} Because the salve randomizes which objects grow each run,
    players develop flexible routing instincts rather than memorizing a fixed solution.
    \item \textbf{Cooperative Potential.} The three-character system naturally supports both solo
    play (switching) and local co-op (one character each), broadening the audience without
    requiring a separate design.
\end{itemize}

\section*{Target Customer}

Fans of platformers and adventure games who enjoy improvisation, character switching, and
environmental spectacle. Players who liked the ``tiny explorer in a big world'' fantasy of games
such as \textit{Pikmin} or \textit{A Bug's Life}. Ages 10 and up; accessible solo, more chaotic
and social in co-op.

\section*{Unique Selling Points}

\begin{itemize}[leftmargin=1.5em, itemsep=2pt]
    \item A craft shop full of ordinary objects reimagined as dramatic physical terrain, providing
    an instantly recognizable yet completely fresh setting.
    \item Three complementary specialist characters whose abilities must be combined to navigate
    any given obstacle, making character switching the primary puzzle mechanic.
    \item A live environmental randomizer (the spreading salve) that ensures no two playthroughs
    produce identical level layouts, adding replayability without extra content.
\end{itemize}

\section*{Competition}

\textit{Pikmin} shares the ``tiny character in an oversized world'' premise but is a real-time
strategy game. \textit{It Takes Two} uses character-ability pairing but is not scale-based.
\textit{Micromancer's Realm} is unique in combining the scale-fantasy setting with a live emergent
environment, a countdown threat, and multi-specialist switching in a single 2D platformer loop.

\section*{Target Hardware}

PC (Windows), implemented in Unreal Engine 5. Mouse-and-keyboard or gamepad supported; local
co-op via split gamepad.

\end{document}
