\documentclass[11pt, letterpaper]{article}
\usepackage[margin=1in]{geometry}
\usepackage{enumitem}
\usepackage{parskip}
\usepackage{titlesec}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{pgfgantt}
\usepackage[hidelinks]{hyperref}

% Section formatting
\titleformat{\section}{\normalfont\large\bfseries}{\thesection}{1em}{}[\titlerule]
\titleformat{\subsection}{\normalfont\normalsize\bfseries}{\thesubsection}{1em}{}
\titleformat{\subsubsection}{\normalfont\normalsize\itshape}{\thesubsubsection}{1em}{}
\titlespacing{\section}{0pt}{12pt}{6pt}
\titlespacing{\subsection}{0pt}{8pt}{4pt}
\titlespacing{\subsubsection}{0pt}{6pt}{2pt}

\begin{document}

% ─── Title Page ──────────────────────────────────────────────────────────────
\begin{titlepage}
    \centering
    \vspace*{2cm}
    {\large University of Pennsylvania}\\[0.5em]
    {\large CIS 5680: Game Design Practicum}\\[3cm]
    {\LARGE \textbf{Game Design Document}}\\[1em]
    {\Huge \textbf{Umbra}}\\[0.5em]
    {\large A Light and Shadow Puzzle Game}\\[3cm]
    {\normalsize Group Members: [Name 1], [Name 2], [Name 3]}\\[1em]
    {\normalsize Instructor: Dr.\ Stephen Lane}\\[1em]
    {\normalsize Spring 2026}
    \vfill
\end{titlepage}

% ─── Table of Contents ───────────────────────────────────────────────────────
\tableofcontents
\newpage

% ─── 1. Executive Summary ────────────────────────────────────────────────────
\section{Executive Summary}

\textbf{Umbra} is a top-down 2D spatial puzzle game built in Unreal Engine~5 for PC
(Windows). Shadows are rendered continuously and update in real time as objects in the
scene move: every light source and blocking object casts live shadow geometry that
physically affects the scene. Players manipulate portable lanterns and pivotable
spotlights---dragging and rotating them freely with the mouse within designer-specified
boundaries---to reshape the shadow pattern across the environment. Blocking objects
(crates, pillars) slide along fixed trajectories to
designer-specified stop positions when the player interacts with them. Shadow regions are
physically solid---a gap becomes crossable only when a shadow-bridge is cast across it,
and displacing that bridge destroys the path. Each level is a self-contained puzzle with
a correct configuration of light sources and blocking objects that opens a walkable path
from start to exit.

Umbra targets fans of clean, mechanic-first puzzle games such as \textit{The Witness},
\textit{Portal}, and \textit{Stephen's Sausage Roll}. There is no combat, no time
pressure, and no randomness---every solution is deterministic and discoverable through
spatial reasoning alone. A single foundational rule---shadows are physically solid---combines
with freely positionable light sources and trajectory-constrained blocking objects to
produce a surprisingly deep puzzle space without requiring the player to learn a large
system. The game is accessible to players aged
12~and~up and designed for pick-up-and-play sessions of 5--15 minutes per level.

Development is structured in three phases across a three-person team: an \emph{Alpha}
prototype delivering one fully playable level with the core shadow mechanic
(due Wed.~Mar.~18); a \emph{Beta} expanding to multiple puzzle levels with
polished art, audio, and a hint system (due Wed.~Mar.~25); and a \emph{Final} release
incorporating a tutorial level, level-select screen, and gameplay/trailer videos (due
Mon.~Mar.~30 and Wed.~Apr.~1). The full prototype is expected to be completed in
approximately five weeks of part-time development.

% ─── 2. Game Design ──────────────────────────────────────────────────────────
\section{Game Design}

    \subsection{High Concept}
    In \textbf{Umbra}, shadows are not merely visual---they are solid, traversable
    terrain. Players manipulate portable lanterns and pivotable spotlights by dragging
    and rotating them with the mouse to cast shadows across the scene. Any region the
    shadow pattern covers becomes physically solid ground; lit regions are open space.
    Blocking objects (crates, pillars) can be slid along fixed trajectories to redirect
    or reshape shadow geometry. Each level is a self-contained spatial puzzle: find the
    correct positions and orientations of all light sources and the correct positions of
    all blocking objects to open a continuous shadow path from start to exit.

    The game is a top-down 2D puzzle built in Unreal Engine~5 for PC (Windows). There is
    no combat, no time pressure, and no randomness. The entire puzzle space emerges from
    one rule: \emph{shadows are solid}.

    \subsection{Design Goals}

        \subsubsection{Main Design Features}

            \paragraph{Player Goals and Objectives}
            The player is an explorer navigating a world where light and shadow are as
            solid as stone. The goal of each self-contained level is to discover the
            correct positions and orientations of light sources and the correct positions
            of blocking objects along their fixed trajectories, casting a shadow pattern
            that opens a walkable path from start to exit. There is no time pressure and no combat; the sole
            challenge is spatial reasoning.

            \paragraph{Main Rules and Procedures}
            \begin{itemize}[leftmargin=1.5em, itemsep=2pt]
                \item Shadows are continuous and computed in real time. Any region covered by a shadow pattern
                      is physically solid and traversable; lit regions are open space.
                      Bridges, platforms, and doorways built from shadow geometry appear
                      or vanish as the shadow pattern changes.
                \item \textbf{Light sources} (lanterns, spotlights) are freely moved or
                      rotated by clicking and dragging with the mouse. Each light source
                      has a designer-specified boundary that limits how far it can travel
                      or rotate. Shadow geometry updates continuously as the player drags.
                \item \textbf{Blocking objects} (crates, pillars) slide along fixed,
                      pre-defined trajectories and snap to designer-specified stop
                      positions when the player clicks and drags them. They cannot be
                      freely placed; each object has exactly the set of positions its
                      trajectory permits.
                \item The level is solved when the player's character reaches the exit
                      through a walkable path that exists in the current shadow
                      configuration.
            \end{itemize}

            \paragraph{Player Resources}
            \begin{itemize}[leftmargin=1.5em, itemsep=2pt]
                \item \textbf{Portable lanterns.} Freely dragged to any position within
                      their movement boundary by clicking and holding with the mouse.
                      Cast a circular point-light shadow pattern that updates in real
                      time as the lantern is repositioned.
                \item \textbf{Wall-mounted spotlights.} Freely rotated about their fixed
                      mount point by clicking and dragging. Emit a directed cone of
                      light; rotating the spotlight continuously reshapes the shadow
                      it casts behind any blocking object in its beam.
                \item \textbf{Blocking objects} (crates, pillars). Slid to any of their
                      trajectory's designated stop positions by clicking and dragging
                      along their fixed path. Reposition them to add, remove, or redirect
                      a shadow cast by a nearby light source.
            \end{itemize}

            \paragraph{Conflicts}
            \begin{itemize}[leftmargin=1.5em, itemsep=2pt]
                \item \textbf{Spatial logic.} The light source positions and blocking object
                      stop positions that solve the puzzle are hidden;
                      players must deduce the correct configuration through observation
                      and reasoning rather than trial and error.
                \item \textbf{Shadow interdependence.} Moving a light source or blocking
                      object continuously reshapes the entire shadow pattern. A position
                      that creates one bridge may simultaneously remove another the player
                      needs. Players must reason about the full shadow layout, not just
                      local effects.
                \item \textbf{Trajectory constraints.} Blocking objects travel only along
                      their fixed paths and stop only at designer-specified positions.
                      The puzzle solution must be achievable within those constraints;
                      players cannot place objects arbitrarily.
            \end{itemize}

        \subsubsection{Appeal}
        Umbra targets fans of pure, mechanic-first puzzle games---players aged 12~and~up
        who enjoy titles such as \textit{The Witness} and \textit{Portal}. A single
        elegant rule (shadows are solid) produces a deep puzzle space accessible to
        newcomers yet satisfying for veterans. Every level has a clean ``aha'' moment
        when the correct light arrangement snaps into place. No combat, timers, or
        randomness---only the satisfaction of spatial reasoning.

        \subsubsection{Look and Feel}
        Umbra uses a \textbf{minimalist silhouette aesthetic} with high contrast between
        lit and shadowed regions. The live shadow boundary shifts continuously as the
        player manipulates light sources or blocking objects, making cause and effect
        immediately visible. Objects are bold geometric silhouettes with no decorative
        detail; light sources are visually distinct from passive blockers. Ambient audio
        is calm and atmospheric, reinforced by crisp sound cues for each interaction.

    \subsection{Worlds, Characters and Story}

        \subsubsection{Back Story}
        An ancient Roman civilization once mastered the art of sculpting light---using
        mirrors, lenses, and sacred flames to build bridges and open sealed vaults across
        their floating island temples. When the empire fell, the islands drifted apart
        and the light paths collapsed. The player is an explorer who has rediscovered one
        of these ruined sanctuaries, and must use the lanterns and spotlights left behind
        to reconstruct the shadow geometry that once connected the fragments and reach the
        vault at the heart of the ruins.

        \subsubsection{Spaces / Worlds}
        All levels are set on the ruins of floating Roman island temples suspended above
        a void. Each level is a single self-contained island platform viewed from a fixed
        top-down perspective. The game ships with four spaces: one tutorial and three
        puzzle levels.

        \begin{itemize}[leftmargin=1.5em, itemsep=4pt]
            \item \textbf{Tutorial --- The Antechamber.} A small, open platform with one
                  lantern and one blocking pillar. Introduces the core mechanic: the player
                  drags the lantern to cast a shadow across a gap, walks across it, then
                  slides the pillar to extend the shadow to the exit. No dead ends; the
                  correct configuration is reached through minimal experimentation.

            \item \textbf{Level 1 --- The Collapsed Bridge.} A medium-sized island with a
                  wide central chasm. One lantern and two blocking pillars are available.
                  The player must position the lantern and arrange the pillars to cast
                  overlapping shadows that span the gap. Introduces the idea that multiple
                  objects must be configured together.

            \item \textbf{Level 2 --- The Torch Hall.} A longer, corridor-like island with
                  two wall-mounted spotlights and two blocking objects. The spotlights must
                  be rotated to project shadows down branching side paths; only one branch
                  leads to the exit. Introduces spotlight rotation and the need to reason
                  about shadow direction.

            \item \textbf{Level 3 --- The Inner Sanctum.} The largest island, combining
                  a lantern, two spotlights, and three blocking objects. The path to the
                  exit requires using all light sources in concert; adjusting one
                  disrupts another. Intended as the capstone challenge that exercises all
                  mechanics introduced in earlier levels.
        \end{itemize}

        \subsubsection{Characters}
        The sole character is \textbf{the Orb}, a Roman ceremonial stone sphere that
        once served as the focal lens at the heart of each island temple. When the empire
        fell, the Orb was scattered across the ruins and lost its light-guiding purpose.
        The player controls the Orb, rolling it through the reconstructed shadow paths to
        return it to the vault at the level's exit. There are no other characters, enemies,
        or NPCs; the Orb navigates the ruins alone.

        \subsubsection{Levels of Difficulty}
        % TODO: Describe how puzzle complexity scales across levels.

    \subsection{Interaction Models}

        \subsubsection{User Interface --- Navigation and Movement}
        % TODO: Describe controls (mouse/keyboard, gamepad), camera perspective,
        %       and how the player moves through the world.

        \subsubsection{Game Play Sequence and Levels}
        % TODO: Describe the flow from level start to completion, including light source
        %       and blocking object interaction, and win/lose conditions.

        \subsubsection{User/Environment --- Obstacles and Props}
        % TODO: Describe interactable objects (lanterns, spotlights, crates,
        %       pillars) and environmental hazards.

        \subsubsection{User/Character}
        % TODO: Describe how the player interacts with the player character
        %       (dragging light sources, sliding blocking objects, etc.).

        \subsubsection{Character/Character}
        % TODO: Describe any character-to-character interactions (if applicable).

    \subsection{Performance and Scoring}

        \subsubsection{State Variables}
        % TODO: List game state variables tracked at runtime (light source positions/
        %       orientations, blocking object stop positions, level completion flag, etc.).

        \subsubsection{Feedback}
        % TODO: Describe visual/audio feedback given to the player (shadow
        %       updates, success/failure cues, hints, etc.).

        \subsubsection{Performance and Progress Metrics}
        % TODO: Describe how player performance is measured and displayed
        %       (move count, time, star rating, level select, etc.).

% ─── 3. Implementation ───────────────────────────────────────────────────────
\section{Implementation}

    \subsection{Design Assumptions}

        \subsubsection{Hardware}
        % TODO: List minimum and recommended hardware requirements (CPU, RAM, GPU).

        \subsubsection{Software}
        % TODO: List software dependencies: Unreal Engine 5 version, OS (Windows),
        %       development tools, version control, etc.

        \subsubsection{Algorithms and Techniques}
        % TODO: Describe key technical approaches: Unreal Engine 5 real-time dynamic
        %       shadow system, continuous shadow geometry as physical collision, mouse
        %       drag interaction with boundary constraints for light sources, fixed
        %       trajectory/snap logic for blocking objects.

    \subsection{Storyboards}
    % TODO: Insert concept sketches or storyboard panels showing key game states
    %       (level start, light source drag, blocking object slide, level completion).

    \subsection{Design Logic}

        \subsubsection{Finite State Machines --- Events / States / Effects}
        % TODO: Provide FSM diagrams or tables for the game's core states:
        %       Idle, Light Source Dragging, Blocking Object Sliding, Level Complete, etc.

        \subsubsection{User Solution / Actions}
        % TODO: Walk through an example puzzle solution step-by-step, mapping
        %       player actions to state changes.

    \subsection{Software Versions}

        \subsubsection{Alpha Version}
        % TODO: List features included in the Alpha (vertical slice):
        %       one playable level, core shadow mechanic, light source drag, blocking
        %       object slide, basic UI.

        \subsubsection{Beta Version}
        % TODO: List additional features for Beta: multiple levels, polished
        %       art/audio, hint system, level select screen.

        \subsubsection{Tutorials}
        % TODO: Describe the in-game tutorial or introductory level that teaches
        %       the shadow mechanic, light source dragging, and blocking object sliding.

% ─── 4. Work Plan ────────────────────────────────────────────────────────────
\section{Work Plan}

    \subsection{Tasks}
    % TODO: Provide a numbered task/subtask breakdown with assigned team members
    %       and estimated durations. Example structure:
    %
    %  1.  Shadow/Light System
    %      1.1 Real-time dynamic shadow rendering (UE5)
    %      1.2 Shadow geometry as physical collision volumes
    %  2.  Level Design
    %      2.1 Puzzle layout design (3 levels for alpha)
    %      2.2 Object placement editor
    %  3.  Player Interaction
    %      3.1 Light source drag (mouse interaction, boundary constraints)
    %      3.2 Blocking object slide (trajectory/snap logic)
    %  4.  UI & HUD
    %      4.1 Level select screen
    %      4.2 Move counter / timer
    %  5.  Art & Audio
    %      5.1 Tile art (lit / shadow)
    %      5.2 Object models
    %      5.3 Sound effects & music
    %  6.  Testing & Polish

    \subsection{Milestones}

        \subsubsection{Minor Milestones}
        % TODO: List internal checkpoints with target dates.

        \subsubsection{Major Milestones}
        % TODO: List the course-defined major milestones:
        %  - Mon. Mar. 2:  Game Design Doc
        %  - Wed. Mar. 18: Alpha Version / Playtest
        %  - Wed. Mar. 25: Beta Version / Playtest
        %  - Mon. Mar. 30: Final Version / Demo
        %  - Wed. Apr. 1:  Game Trailer and Gameplay Videos

    \subsection{Development Schedule}
    % TODO: Insert a Gantt chart mapping Work Plan tasks to calendar weeks.
    %       The pgfgantt package is loaded above for this purpose.

\end{document}
